---
title: "Davidson County Voter Analytics | Masheke Analytics"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: flatly
    social: menu
    source_code: false
---

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(tidyr)
library(sf)
library(tigris)
library(leaflet)
library(plotly)
library(DT)
library(htmltools)
library(scales)

options(tigris_use_cache = TRUE)
```

```{r data-processing, include=FALSE}

dcvoter <- read.table(
  "C:/Users/richa/OneDrive/Desktop/ResearchData/PoliticalData/dcvoter.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

status_2 <- dcvoter %>% 
  filter(voter_status_desc == "ACTIVE") %>% 
  select(voter_reg_num, ncid, zip_code, party_cd, age_at_year_end,
         gender_code,race_code,precinct_desc, ward_desc,cong_dist_abbrv, mail_city)

voter_counts <- status_2 %>%
  count(zip_code, party_cd, name = "n") %>%
  filter(!is.na(zip_code), zip_code != "NA") %>%
  mutate(zip_code = as.character(zip_code)) %>%
  pivot_wider(
    names_from = party_cd,
    values_from = n,
    values_fill = list(n = 0)
  )

needed_cols <- c("DEM", "REP", "UNA", "LIB", "GRE")

for (col in needed_cols) {
  if (!col %in% names(voter_counts)) {
    voter_counts[[col]] <- 0
  }
}

voter_counts <- voter_counts %>%
  mutate(
    total_active_voters = DEM + REP + UNA + LIB + GRE,
    dem_share = DEM / total_active_voters,
    rep_share = REP / total_active_voters,
    una_share = UNA / total_active_voters,
    lib_share = LIB / total_active_voters,
    gre_share = GRE / total_active_voters,
    leading_party = case_when(
      DEM >= REP & DEM >= UNA & DEM >= LIB & DEM >= GRE ~ "DEM",
      REP >= DEM & REP >= UNA & REP >= LIB & REP >= GRE ~ "REP",
      UNA >= DEM & UNA >= REP & UNA >= LIB & UNA >= GRE ~ "UNA",
      LIB >= DEM & LIB >= REP & LIB >= UNA & LIB >= GRE ~ "LIB",
      GRE >= DEM & GRE >= REP & GRE >= UNA & GRE >= LIB ~ "GRE",
      TRUE ~ "Other"
    )
  ) %>%
  rowwise() %>%
  mutate(
    party_gap = {
      vals <- sort(c(DEM, REP, UNA, LIB, GRE), decreasing = TRUE)
      (vals[1] - vals[2]) / total_active_voters
    }
  ) %>%
  ungroup()

nc_zctas <- zctas(
  state = "NC",
  year = 2000,
  cb = TRUE,
  class = "sf"
)

davidson_boundary <- counties(
  state = "NC",
  cb = TRUE,
  class = "sf"
) %>%
  filter(NAME == "Davidson")

davidson_map_data <- nc_zctas %>%
  st_filter(davidson_boundary)

final_map_data <- davidson_map_data %>%
  inner_join(voter_counts, by = c("ZCTA" = "zip_code"))

total_v <- sum(final_map_data$total_active_voters, na.rm = TRUE)

p_totals <- colSums(
  st_drop_geometry(final_map_data)[, c("DEM", "REP", "UNA", "LIB", "GRE")],
  na.rm = TRUE
)

p_df <- data.frame(
  Party = names(p_totals),
  Count = as.numeric(p_totals)
)
```

Voter Overview
=====================================

Column {data-width=250 .sidebar}
-------------------------------------

### Executive Summary

Based on current active voter registration data for Davidson County:

**Party Dominance:** Republican registration leads across much of the county.

**Key Hubs:** ZIP codes 27295 and 27360 represent major concentrations of active voters.

**Unaffiliated Growth:** Unaffiliated voters are a critical segment and may influence local election outcomes.

<br>

**Produced by Masheke Analytics**

Column {data-width=750}
-------------------------------------

### Interactive Voter Density Map

```{r voter-map}
pal <- leaflet::colorNumeric(
  palette = "YlOrRd",
  domain = final_map_data$total_active_voters
)

leaflet(data = final_map_data) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    fillColor = ~pal(total_active_voters),
    weight = 1,
    color = "white",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 3,
      color = "#666",
      bringToFront = TRUE
    ),
    label = lapply(1:nrow(final_map_data), function(i) {
      HTML(paste0(
        "<b>ZIP Code: </b>", final_map_data$ZCTA[i], "<br>",
        "<b>Total Active Voters: </b>", format(final_map_data$total_active_voters[i], big.mark = ","), "<br>",
        "<b>DEM: </b>", format(final_map_data$DEM[i], big.mark = ","), "<br>",
        "<b>REP: </b>", format(final_map_data$REP[i], big.mark = ","), "<br>",
        "<b>UNA: </b>", format(final_map_data$UNA[i], big.mark = ","), "<br>",
        "<b>LIB: </b>", format(final_map_data$LIB[i], big.mark = ","), "<br>",
        "<b>GRE: </b>", format(final_map_data$GRE[i], big.mark = ",")
      ))
    }),
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "13px",
      direction = "auto"
    )
  ) %>%
  addLegend(
    pal = pal,
    values = ~total_active_voters,
    title = "Active Voters",
    position = "bottomright"
  )
```

Column {data-width=350}
-------------------------------------

### Total Active Voters

```{r total-active-voters}
valueBox(
  format(total_v, big.mark = ","),
  icon = "fa-users",
  color = "info"
)
```

### County-Wide Party Distribution

```{r party-distribution}
plot_ly(
  p_df,
  x = ~Party,
  y = ~Count,
  type = "bar"
) %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Total Count")
  )
```

### Searchable Data by ZIP

```{r zip-table}
voter_counts %>%
  select(zip_code, total_active_voters, DEM, REP, UNA, LIB, GRE) %>%
  arrange(desc(total_active_voters)) %>%
  datatable(
    options = list(pageLength = 6, dom = "tp"),
    rownames = FALSE
  )
```

Voter Deep Dive
=====================================

Column {data-width=500}
-------------------------------------

### Party Share by ZIP

```{r party-share}
voter_counts %>%
  select(zip_code, dem_share, rep_share, una_share) %>%
  pivot_longer(
    cols = c(dem_share, rep_share, una_share),
    names_to = "party",
    values_to = "share"
  ) %>%
  mutate(
    party = recode(
      party,
      dem_share = "DEM",
      rep_share = "REP",
      una_share = "UNA"
    )
  ) %>%
  plot_ly(
    x = ~zip_code,
    y = ~share,
    color = ~party,
    type = "bar"
  ) %>%
  layout(
    barmode = "stack",
    yaxis = list(title = "Share of Active Voters", tickformat = ".0%"),
    xaxis = list(title = "ZIP Code")
  )
```

### Unaffiliated Voter Concentration

```{r una-concentration}
voter_counts %>%
  arrange(desc(UNA)) %>%
  mutate(una_share_label = sprintf("%.1f%%", una_share * 100)) %>%
  select(zip_code, total_active_voters, UNA, una_share_label) %>%
  datatable(
    colnames = c("ZIP Code", "Total Active Voters", "UNA Count", "UNA Share"),
    options = list(pageLength = 8, dom = "tp"),
    rownames = FALSE
  )
```

Column {data-width=500}
-------------------------------------

### Competitive Balance by ZIP

```{r competitive-balance}
voter_counts %>%
  arrange(party_gap) %>%
  mutate(
    party_gap_label = sprintf("%.1f%%", party_gap * 100)
  ) %>%
  plot_ly(
    x = ~reorder(zip_code, party_gap),
    y = ~party_gap,
    type = "bar",
    text = ~party_gap_label,
    textposition = "inside",
    insidetextanchor = "middle",
    textangle = 0,
    hovertemplate =
      paste(
        "<b>ZIP:</b> %{x}<br>",
        "<b>Leading Party:</b> %{customdata}<br>",
        "<b>Gap:</b> %{text}<extra></extra>"
      ),
    customdata = ~leading_party
  ) %>%
  layout(
    xaxis = list(title = "ZIP Code"),
    yaxis = list(
      title = "Gap Between Top Two Parties",
      tickformat = ".0%"
    ),
    showlegend = FALSE
  )

```

### ZIP-Level Deep Dive Table

```{r deep-dive-table}
voter_counts %>%
  mutate(
  dem_share = sprintf("%.1f%%", dem_share * 100),
  rep_share = sprintf("%.1f%%", rep_share * 100),
  una_share = sprintf("%.1f%%", una_share * 100),
  party_gap = sprintf("%.1f%%", party_gap * 100)
) %>%
  select(
    zip_code,
    total_active_voters,
    DEM,
    REP,
    UNA,
    dem_share,
    rep_share,
    una_share,
    leading_party,
    party_gap
  ) %>%
  arrange(desc(total_active_voters)) %>%
  datatable(
    colnames = c(
      "ZIP Code",
      "Total Active Voters",
      "DEM",
      "REP",
      "UNA",
      "DEM %",
      "REP %",
      "UNA %",
      "Leading Party",
      "Party Gap"
    ),
    options = list(pageLength = 8, dom = "tp"),
    rownames = FALSE
  )
```

Demographic Intelligence
=====================================

```{r demographic-data, include=FALSE}
voter_demo <- status_2 %>%
  mutate(
    age_at_year_end = as.numeric(age_at_year_end),
    zip_code = as.character(zip_code),
    age_group = case_when(
      age_at_year_end < 25 ~ "18-24",
      age_at_year_end < 35 ~ "25-34",
      age_at_year_end < 45 ~ "35-44",
      age_at_year_end < 55 ~ "45-54",
      age_at_year_end < 65 ~ "55-64",
      age_at_year_end >= 65 ~ "65+",
      TRUE ~ "Unknown"
    ),
    gender_label = case_when(
      gender_code == "F" ~ "Female",
      gender_code == "M" ~ "Male",
      TRUE ~ "Unknown"
    ),
    race_label = case_when(
      race_code == "W" ~ "White",
      race_code == "B" ~ "Black",
      race_code == "A" ~ "Asian",
      race_code == "I" ~ "American Indian",
      race_code == "M" ~ "Multiracial",
      race_code == "O" ~ "Other",
      TRUE ~ "Unknown"
    )
  )

age_party <- voter_demo %>%
  filter(party_cd %in% c("DEM", "REP", "UNA")) %>%
  count(age_group, party_cd)

gender_party <- voter_demo %>%
  filter(party_cd %in% c("DEM", "REP", "UNA")) %>%
  count(party_cd, gender_label)

race_party <- voter_demo %>%
  filter(party_cd %in% c("DEM", "REP", "UNA")) %>%
  count(party_cd, race_label)

zip_age <- voter_demo %>%
  group_by(zip_code) %>%
  summarise(
    median_age = median(age_at_year_end, na.rm = TRUE),
    active_voters = n(),
    .groups = "drop"
  ) %>%
  filter(!is.na(zip_code), zip_code != "NA") %>%
  arrange(desc(active_voters))
```

Column {data-width=250 .sidebar}
-------------------------------------

### Key Demographic Insights

This page profiles Davidson County active voters by age, gender, race, and ZIP-level age patterns.

These views help identify:

**Age Structure:** Which parties have older or younger voter bases.

**Gender Mix:** How male and female voters are distributed across party registration.

**Race Composition:** How race varies across major party groups.

**ZIP Age Profile:** Which ZIP codes have older or younger registered voter populations.

Column {data-width=500}
-------------------------------------

### Age Distribution by Party

```{r age-party-chart}
plot_ly(
  age_party,
  x = ~age_group,
  y = ~n,
  color = ~party_cd,
  type = "bar"
) %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Age Group"),
    yaxis = list(title = "Active Voters")
  )
```

### Gender Composition by Party

```{r gender-party-chart}
plot_ly(
  gender_party,
  x = ~party_cd,
  y = ~n,
  color = ~gender_label,
  type = "bar"
) %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Party"),
    yaxis = list(title = "Active Voters")
  )
```

Column {data-width=500}
-------------------------------------

### Race Composition by Party

```{r race-party-chart}
plot_ly(
  race_party,
  x = ~party_cd,
  y = ~n,
  color = ~race_label,
  type = "bar"
) %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Party"),
    yaxis = list(title = "Active Voters")
  )
```

### Median Age by ZIP

```{r median-age-zip-chart}
plot_ly(
  zip_age,
  x = ~reorder(zip_code, median_age),
  y = ~median_age,
  type = "bar",
  text = ~round(median_age, 1),
  textposition = "auto"
) %>%
  layout(
    xaxis = list(title = "ZIP Code"),
    yaxis = list(title = "Median Age")
  )
```

### ZIP Demographic Table

```{r zip-demo-table}
zip_age %>%
  datatable(
    colnames = c("ZIP Code", "Median Age", "Active Voters"),
    options = list(pageLength = 8, dom = "tp"),
    rownames = FALSE
  )
```


Electoral Geography
=====================================

Column {data-width=500}
-------------------------------------

### Top Precincts by Active Voters

```{r top-precincts}
precinct_counts <- voter_demo %>%
  filter(!is.na(precinct_desc), precinct_desc != "") %>%
  count(precinct_desc, party_cd, name = "n") %>%
  pivot_wider(
    names_from = party_cd,
    values_from = n,
    values_fill = list(n = 0)
  )

needed_cols_geo <- c("DEM", "REP", "UNA")

for (col in needed_cols_geo) {
  if (!col %in% names(precinct_counts)) {
    precinct_counts[[col]] <- 0
  }
}

precinct_counts <- precinct_counts %>%
  mutate(
    total_active_voters = DEM + REP + UNA
  ) %>%
  arrange(desc(total_active_voters))

precinct_counts %>%
  select(precinct_desc, total_active_voters, DEM, REP, UNA) %>%
  head(15) %>%
  datatable(
    colnames = c("Precinct", "Total Active Voters", "DEM", "REP", "UNA"),
    options = list(pageLength = 8, dom = "tp"),
    rownames = FALSE
  )
```

### Precinct Party Composition

```{r precinct-party-composition}
precinct_counts %>%
  head(12) %>%
  select(precinct_desc, DEM, REP, UNA) %>%
  pivot_longer(
    cols = c(DEM, REP, UNA),
    names_to = "party",
    values_to = "voters"
  ) %>%
  plot_ly(
    x = ~reorder(precinct_desc, voters),
    y = ~voters,
    color = ~party,
    type = "bar"
  ) %>%
  layout(
    barmode = "stack",
    xaxis = list(title = "Precinct"),
    yaxis = list(title = "Active Voters")
  )
```

Column {data-width=500}
-------------------------------------

### Most Competitive Precincts

```{r competitive-precincts}
precinct_comp <- precinct_counts %>%
  filter(total_active_voters > 0) %>%
  rowwise() %>%
  mutate(
    leading_party = case_when(
      DEM >= REP & DEM >= UNA ~ "DEM",
      REP >= DEM & REP >= UNA ~ "REP",
      UNA >= DEM & UNA >= REP ~ "UNA",
      TRUE ~ "Other"
    ),
    party_gap = {
      vals <- sort(c(DEM, REP, UNA), decreasing = TRUE)
      (vals[1] - vals[2]) / total_active_voters
    },
    party_gap_label = sprintf("%.1f%%", party_gap * 100)
  ) %>%
  ungroup() %>%
  arrange(party_gap)

precinct_comp %>%
  head(15) %>%
  plot_ly(
    x = ~reorder(precinct_desc, party_gap),
    y = ~party_gap,
    type = "bar",
    text = ~party_gap_label,
    textposition = "outside",
    hovertemplate = paste(
      "<b>Precinct:</b> %{x}<br>",
      "<b>Leading Party:</b> %{customdata}<br>",
      "<b>Gap:</b> %{text}<extra></extra>"
    ),
    customdata = ~leading_party
  ) %>%
  layout(
    xaxis = list(title = ""),
    yaxis = list(title = "Gap Between Top Two Parties", tickformat = ".0%"),
    showlegend = FALSE,
    margin = list(t = 60)
  )
```

### Voter Geography by Mail

```{r city-voter-summary}
city_counts <- voter_demo %>%
  filter(!is.na(mail_city), mail_city != "") %>%
  count(mail_city, party_cd, name = "n") %>%
  pivot_wider(
    names_from = party_cd,
    values_from = n,
    values_fill = list(n = 0)
  )

for (col in needed_cols_geo) {
  if (!col %in% names(city_counts)) {
    city_counts[[col]] <- 0
  }
}

city_counts <- city_counts %>%
  mutate(
    total_active_voters = DEM + REP + UNA
  ) %>%
  arrange(desc(total_active_voters))

city_counts %>%
  select(mail_city, total_active_voters, DEM, REP, UNA) %>%
  head(15) %>%
  datatable(
    colnames = c("City", "Total Active Voters", "DEM", "REP", "UNA"),
    options = list(pageLength = 8, dom = "tp"),
    rownames = FALSE
  )
```

```{r}

dcvoter_nc <- read.table(
  "C:/Users/richa/OneDrive/Desktop/ResearchData/PoliticalData/ncvoter29.txt",
  header = TRUE,
  sep = "\t",
  stringsAsFactors = FALSE
)

```

```{r}

system("git status")
system("git add index.Rmd index.html")
system('git commit -m "Update Davidson County voter analytics report"')
system("git push")

```




```







